---
output:
  html_document:
    toc: true
    number_sections: true
    highlight: tango
    theme: cerulean
---

**Extensive form games**


```{r include=FALSE, cache=FALSE}
rm(list=ls(all=TRUE))
library(knitr)
```

# Getting Going   
First open R or Rstudio and install the hop package from github (make sure you have devtools installed to be able to run the below command): 

```{r eval=FALSE}
devtools::install_github("macartan/hop")
```

Then load it up:
```{r}
library(hop)
```

# Game trees
The ```gt_tree``` function graphs game trees given a full specification of an extensive form game with complete information. To define an extensive form game we need to define:

* $H$: a set of histories -- that is all the possible paths that can be followed
* $P$: a player function that says what player moves at what point in each history
* $U$: Utilities awarded to players at the end of each history

## Illustration 1; Sequential Chicken

Here is a simple example. Lets imagine a Chicken game played out over time. H, P and U, are defined as follows:


```{r}
H <- matrix(c("0", "0", "0", "0", "C", "C", "D", "D", "C", "D",  "C", "D"),4)
P <- matrix(c(rep(1,4), rep(2,4)),4)
U <- matrix(c(1,0,2,-1,1,2, 0, -1), 4)
```

Then the game tree looks like this:
  
```{r, fig.align = "center"}
gt_tree(H,U,P, title = "Sequential Chicken", solution = TRUE)
```

Note that `gt_tree` interprets the set of histories as a tree; it figures out the branches based on when the histories depart from each other and the nodes from the player function. The underlying data looks like this: 

History:
```{r}
rownames(H) <- c("H1", "H2", "H3", "H4")
kable(H, col.names = c("T0", "T1", "T2"))
```

Player Function:
```{r}
rownames(P) <- c("H1", "H2", "H3", "H4")
kable(P, col.names = c("T1", "T2"))
```

Utilities:
```{r}
rownames(U) <- c("H1", "H2", "H3", "H4")
kable(U, col.names =c("U1", "U2"))
```

## Illustration 2: The Hostage Game
The introductory text describes a hostage game in which options depend on past history.

Here is the graph with the solution:
```{r}
gt_tree(
      H = matrix(c("Take hostages", "Take hostages", "Don't take", "Pay Ransom", "Don't pay", ""),3),
      U = matrix(c(1,-1,0,-1,-2,0), 3),
      P = matrix(c(rep(1,3), c(2,2,1)),3),
      solution=TRUE, 
      player.names=c("Militants", "Gov")
      )
```

# More complex games
The function can allow much more complex games; the key chllenge is to be able to write out all the histories. Somestimes these histories can be generated in an automated way to make families of more complex games. 

```{r}
centipede <- function(n, ...){
		H <- matrix("C", n+1, n+1)
		H[lower.tri(H)] <- "D"
		H <- H[,n:1]
		U <- cbind(U1 = ((n+1):1) + 2*((n+1):1)%%2 - 2,
		          U2 = ((n+1):1) - 2*((n+1):1)%%2 + 1)
		P <-  t(replicate(n+1, rev(rep(1:2, length = n+1))))
		P[lower.tri(P)] <- t(P)[lower.tri(P)]
		P <- t(P)[,(n+1):1]
		gt_tree(H,U,P, ...)
		}
```

Now use this to graph a short and a long centipede game: 
```{r}
centipede(n = 3, solution = TRUE)		
centipede(n = 8, solution = TRUE, textsize=.75)		
```

# Adding and forcing solutions
You can add or remove solutions using the solution argument. The sequential chicken without a solution is drawn like this:

```{r, fig.align = "center"}
gt_tree(H,U,P, title = "Sequential Chicken", solution = FALSE)
```

The solution will however generally not appear for a non-generic game:

```{r, fig.align = "center"}
U[1,2] <- 0
gt_tree(H,U,P, title = "Non-generic", solution = TRUE)
```

Unless forced:
```{r, fig.align = "center"}
gt_tree(H,U,P, title = "Non-generic", solution = TRUE, force_solution = TRUE)
```

This can however have undesireable consequences:
```{r, fig.align = "center"}
U[1,2] <- 2
gt_tree(H,U,P, title = "Non-generic", solution = TRUE, force_solution = TRUE)
```


# Formatting
You can do a reasonable amount of formatting in terms of specifying sizes and colors. eg: 
```{r, fig.align = "center"}
gt_tree(H,U,P, 
  title = "Sequential Chicken",
  force_solution = TRUE,
  player.names=c("Jack", "Jill"),
  textsize  = 1,
  titlesize = 2,
  utextsize = 1.25,
  btextsize = 1.5,
  slopetext = FALSE,

  branchcol = "orange",
  actioncol = "black",
  titlecol  = "grey",
  playercol = "red",
  solutioncol = "darkorange"
        )
```

# Limitations
Besides being in alpha stage the main limitations of this function the fact that it cannot solve when there are multiple equilibria and payoff indifferences; and it cannot solve games of incomplete information. Even still it can be used to graph these and then then information sets and solutions can be superimposed.  
