---
title: 'Political Games'
output:
  html_document:
    highlight: tango
    number_sections: yes
    theme: cerulean
---

*Political Games* (formally known as *Hell is Other People*) presents 49 major results from formal theory that give insights into political bahavior. Many of the results are  illustrated using simple graphs or payoff matrices produced in [R](https://www.r-project.org/).

You can recreate or modify most of these figures using the R package ```hop```. Here's how along with some examples. 

```{r include=FALSE, cache=FALSE}
rm(list=ls(all=TRUE))
```

# Getting Going  
First open R or Rstudio and install the hop package from github:

```{r eval=FALSE}
devtools::install_github("macartan/hop")
```

Then load it up:
```{r}
library(hop)
```

# Payoff matrices
The ```gt_bimatrix``` takes a payoff matrix and turns it into a nice figure. It can also mark best response functions as arrows and mark the (pure strategy) equilibriums as a star.

For example, the prisoner's dilemma matrix looks like this:
```{r}
PD <- matrix(c(2,4,1,3),2,2)
```

Once defined you can plot it like this:

```{r}
gt_bimatrix(PD, nash = FALSE, arrow1 = FALSE,  tscale = .75)
```

And make it nicer like this:

```{r}
gt_bimatrix(PD, nash = FALSE, arrow1 = FALSE, P1 = "Jack", P2 = "Jill",  labels1=c("Cooperate","Defect"), srt1=90, tscale = .5)
```

You can also add in best response arrows that show the incentives to change strategies;  like this:

```{r}
gt_bimatrix(PD, nash = FALSE, arrow1 = TRUE, P1 = "Jack", P2 = "Jill",  labels1=c("Cooperate","Defect"), srt1=90, tscale = .5)
```

And add in a star to mark the Nash equilibria:

```{r}
gt_bimatrix(PD, P1 = "Jack", P2 = "Jill",  labels1=c("Cooperate","Defect"), srt1=90, tscale = .5)
```

These are useful especially for larger, more complex games:

```{r}
Jack <- matrix(10*c(1:10) - (1:10)^2,2)
Jill <- matrix(10*c(1:10) - (1:10)^2,2)
gt_bimatrix(X=Jack, Y=Jill, P1 = "Jack", P2 = "Jill",  tscale = .5)
```


Of for quickly comparing families of games:

```{r}
par(mfrow=c(2,2))
gt_bimatrix(X = matrix(c(2,4,1,3),2,2), tscale = .4, main = "PD")
gt_bimatrix(X = matrix(c(3,4,2,1),2,2), tscale = .4, main = "Chicken")
gt_bimatrix(X = matrix(c(4,2,1,3),2,2), tscale = .4, main = "Assurance")
gt_bimatrix(X = matrix(c(1,4,3,2),2,2), Y = matrix(c(2,3,4,1),2,2), 
            tscale = .4, labels1 = c("Boxing", "Ballet"), srt1=90, 
            main = "Battle of the Sexes")
```

